Performance Evaluation of Big Data Processing Strategies for Neuroimaging

### 2018-12-14 paper.tex

\begin{abstract}
We benchmark in-memory, computing, data locality and lazy Big Data processing 
strategies on typical neuroimaging processing use cases represented by the BigBrain dataset. We use Dell 
EMC's Top-500 cluster, as well as Apache Spark and Nipype as 
representative processing engines. We model Big Data thresholds by 
comparing the data/compute ratio of the application to the disk and 
shared file system bandwidths reported to the number of concurrent 
processes in a node and across the cluster. This model acknowledges the 
fact that page caching provided by the Linux kernel is critical to the 
performance of Big Data applications. Results show that in-memory 
computing alone speeds-up executions by a factor of up to 1.6, while 
combined with data locality this factor reaches 5.3. Such important 
speed-up values are likely to be observed on typical image processing 
operations performed on images of size larger than 75GB. A ballpark 
speculation from our model showed that in-memory computing alone would 
not speed-up current 
functional MRI analyses, but that it would when coupled with data 
locality, for a number of concurrently  processed subjects in the order 
of 200. In addition, we observe that emulating in-memory computing 
using in-memory file systems (tmpfs) does not reach the performance of 
an in-memory engine, presumably due to swapping to disk and the lack of data cleanup.
We conclude that Big Data processing strategies are 
worth developing for neuroimaging 
applications. 
\end{abstract}

@@

Big Data processing engines reduce data-related overheads through 
use of concepts such as data locality, in-memory computing, and lazy evaluation.
Data locality-aware scheduling, popularized by the MapReduce~\cite{dean2008mapreduce} 
framework, runs tasks to the nodes holding the data rather than 
performing costly data transfers across the network. In-memory 
computing, adopted by MapReduce's successor Apache 
Spark~\cite{zaharia2016apache}, stores data in memory whenever 
possible, reducing costly disk I/Os between pipeline steps. 
Lazy evaluation, also used in Spark, means that results are computed only 
when necessary to enable further performance optimizations. Frameworks 
such as
MapReduce and Spark have become mainstream tools for data analytics, 
although many others, such as 
Dask~\cite{rocklin2015dask}, are emerging. 
Meanwhile, several scientific 
domains including bioinformatics, physics or astronomy, have entered 
the Big Data era due to increasing data volumes and variety. 
However, the adoption of Big Data engines for scientific data analysis 
remains limited, perhaps due to the widespread availability of 
scientific processing engines such as Pegasus~\cite{deelman2005pegasus} or
Taverna~\cite{oinn2004taverna}, and the adaptations required in Big 
Data processing engines for scientific computing.

Scientific applications differ from the typical Big Data use 
cases, which might explain the remaining gap between Big Data and 
scientific engines. While Big Data applications mostly target text 
processing (e.g. web search, frequent pattern mining, recommender 
systems~\cite{leskovec2014mining}) implemented in consistent software 
libraries, scientific applications often involve 
binary data such as images and signals, processed by command-line tools 
using a mix of programming languages (C, Fortran, Python, shell 
scripts). Infrastructure-wise, Big Data applications commonly run on 
clouds or dedicated commodity clusters with locality-aware file systems 
such as the Hadoop Distributed File System 
(HDFS~\cite{shvachko2010hadoop}), whereas scientific applications are 
usually deployed on large, shared clusters where data is moved between
data nodes and compute nodes using shared file systems such 
as Lustre~\cite{schwan2003lustre}. Such differences in applications and 
infrastructures have important consequences. To 
mention only one, in-memory computing requires instrumentation to be 
applied to command-line tools.

Technological advances of the past decade, in particular page caching 
in the Linux kernel~\cite{love2010linux}, in-memory file systems 
(tmpfs) and memory-mapped files might also 
explain the lack of adoption of Big Data engines for scientific 
applications.
In-memory computing would then be a feature provided by 
the operating system rather than by the engine itself. The frontier 
between these two components is blurred and needs to be clarified.

@@

Our study focuses on performance. We intentionally do not 
compare Big Data and scientific data processing engines on the grounds 
of workflow language expressivity, fault-tolerance, provenance capture 
and representation, portability or reproducibility which are otherwise 
critical concerns, addressed for instance in~\cite{samba}. Besides, our 
study of performance focuses on the impact of data writes and 
transfers. It purposely leaves out task scheduling to computing 
resources, to focus on the understanding of 
data writes and movement. Task scheduling will be part of our 
discussion though.

@@

\todo{Perhaps talk about modeling and simulation here.}

@@

An RDD is an immutable parallel data structure that achieves fault-tolerance 
through the concept of lineage~\cite{zaharia2010spark}. Rather than permitting
fine-grained transformations, only coarse-grained transformations (applying to
many elements in the RDD), can be applied, thereby making it simple to keep a 
log of how the data was modified. This log, known as the lineage, is used
to reproduce any lost data.

Two types of operations can be performed on RDDs:
transformations and actions. Applying a transformation to an RDD produces a new
child RDD through a narrow or wide dependency. A narrow dependency signifies 
that the child is only dependent on a single parent partition, whereas a child 
RDD is dependent on all parent partitions in a wide dependency. Examples of 
Spark transformations include map, filter and join. To materialize an RDD, an
action must be performed, such as a reduce or a collect. \todo{Lazy evaluation should be mentioned here, or elsewhere in the paragraph.
It is currently not described.}

All actions and wide dependencies require a shuffle -- Spark's most costly
operation. Every shuffle begins with each map task saving its data to local
files to be fetched by the shuffle operation. The shuffle operation then 
redistributes the data across the partitions as requested. A shuffle marks a 
stage boundary in Spark. In other words, a shuffle will not begin until all 
dependent map tasks \todo{shouldn't it be 'actions'?} have completed.

Spark will spill RDD data to disk whenever the complete dataset does not fit in
memory. Moreover, as Spark transformations generate new RDDs and numerous 
transformations may occur in a single application, Spark implements a Least-Recently Used (LRU) eviction
policy. If an evicted RDD needs to be reused, Spark will recompute it using the
lineage data collected. It therefore becomes of great importance to persist to
memory and/or 
disk RDDs which will be reused throughout the application, particularly if the
recomputation is costly. 
%mention results in thunder paper

Data locality in Spark is achieved through the  scheduling of task to partitions which have the
data loaded in memory. If the data is instead stored on HDFS, the scheduler will
assign it to one of the preferred locations specified by HDFS. Spark's scheduler
utilizes delay scheduling to optimize fairness and locality for all tasks.
%improve paragraph

@@

Executing Spark applications on HPC systems with Spark-unaware schedulers may
be inefficient. The amount of resources requested by Spark may impede 
Spark-cluster scheduling time. Using pilot-scheduling strategies to add 
nodes to the Spark cluster as they are allocated by the underlying 
schedulers may speedup allocation and overall processing time 
\todo{Cite Shantenu's paper here}. This, however, is not studied in the 
current paper.

@@

Nipype is a popular Python neuroimaging processing engine. It aims at being a 
solution for easily creating reproducible neuroimaging workflows. Although Nipype
does not employ any Big Data processing strategies, it provides access to numerous
plugins to schedulers found in most clusters readily available to researchers, 
such as SGE, PBS, SLURM and HTCondor\todo{expand acronyms}. It also includes its own scheduler, MultiProc,
for parallel processing on single nodes. Nipype also includes many built-in 
interfaces to commonly used neuroimaging tools to be incorporated within the 
workflows. Nipype's ability to easily parallelize workflows in researcher-available 
cluster setups, capture detailed provenance information necessary for reproducibility,
and allow users to easily integrate existing neuroimaging tools, make it preferable
to existing Big Data solutions, which would necessitate modification to achieve this.

@@

In order for Nipype to operate as intended, a filesystem shared by all the nodes 
is required. However, it is still possible to save to a non-shared local filesystem, 
but it may come at the expense of fault-tolerance \todo{explain why}. Moreover, the user will need to
ensure that the files are appropriately directed to the nodes that require them as
there is no guarantee of data locality in Nipype.

@@

\subsection{Simulation} % is it included? 0.5 page

\todo{Remove if not included}

@@

 All experiments were executed on  
 \href{https://www.dellemc.com/resources/en-us/asset/sales-documents/products/storage/h16221-hpc-lab-brochure.pdf}{Dell EMC's Zenith 
 cluster}, a Top-500 machine in the Dell EMC HPC Innovation Lab, running 
 SLURM. For the Spark experiments, a Spark cluster was started atop our 
 SLURM allocation through access to 16 dedicated nodes. Each compute 
 node ran on RHEL 7.4, was 
 equipped with 40 Intel Xeon Gold 6148 2.4GHz CPUs and had
 187GB of memory, 94GB of tmpfs and a Dell 120GB M.2 SATA SSD as local disk. 
 A 698TB Lustre server was accessible on each node through a 100Gb/s 
 Infiniband network. The apparent write bandwidth of the local disk, 
 Lustre file system and tmpfs were measured by sequentially writing various numbers
 of image blocks containing random intensities, to avoid 
 caching effects (measure\_bandwidth.py script in folder 
 benchmark\_scripts). They are reported in Table~\ref{table:bdwdths}.

@@

% Incrementation (avoid cache effects in binarization)
In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consisted exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the processing engines. As incrementing the images is a relatively quick 
process, we added a sleep delay to the incrementation stage in order to 
increase task duration as required. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the increment chunks would be incremented again and saved to the same storage. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks would be reads from Lustre and the final 
output chunks would be saved to Lustre. We chose to perform our initial 
reads and final writes on Lustre as a multi-tenant clusters' compute node's 
local storage, if 
available, is generally temporary and only available for the duration of the 
allocation.

@@

We conducted 4 experiments in which we varied (1) the number of 
iterations in the application ($n$ in 
Algorithm~\ref{alg:incrementation}), (2) the task duration ($x$), (3) 
the chunk size for a constant image size, (4) the total image size. To 
evaluate the page-cache model, experiment conditions fell in different 
regions of Equations~\ref{eq:page-cache-inequality} 
and~\ref{eq:page-cache-sharedfs}, as summarized in 
Table~\ref{table:experiments}. Among the 16 nodes available, 1 was 
dedicated to the Spark master and driver, and the remaining 15 were 
used as compute nodes. Since data-locality is not normally preserved in 
Nipype (a new SLURM allocation is requested for each processed task), 
we instrumented Nipype to ensure data locality. That is, for each chunk 
partition, we requested a SLURM allocation to process the entire 
pipeline in parallel, using Nipype's MultiProc scheduler, on a given 
node. This was possible as no communication was required between the 
processed chunks.

@@

As a third incrementation experiment, we were interested in the effects of 
chunk size on Big Data strategies. Naturally, greater chunk size signifies 
a decrease in parallelization, however, it also means an increase in 
sequential I/O (increased $\Delta/\Gamma$ and $\delta/\gamma$). For 
this experiment we partitioned the complete BigBrain image into 30, 125 
and 750 chunks, corresponding to $\gamma$ values of 2, 9 and 25 
respectively. While Spark attempted to load-balance the data, it used 
up only 25 of the 40 cores for 750 chunks. In contrast, Nipype tried to 
use up as many cores as possible. To ensure both engines were 
processing the same amount of data concurrently (same $\gamma$ value), 
we fixed the amount of cores used by Nipype to 25. Unlike the previous 
experiment, the D/C ratio was kept static at 178.6MB/s, however, this 
ratio ensured that different regions of the inequality were reached 
depending on amount of parallelization. Unfortunately, 30 chunks could 
no be processed using in-memory Spark as its individual chunk size of 
3GB exceeded the 2GB partition size limit imposed by Spark. The number of iterations was
fixed to 10, and the task duration was adjusted so that C remained constant at 4,400s.

@@

Fig.~\ref{fig:iterations} shows the difference between the different 
filesystem choices given the number of iterations. At 1 iteration, all 
filesystems behave the same, although the application was writing faster than the page cache flushed.
This is explained by the fact that the application wasn't running long enough to saturate the page 
cache (transient phase). As the total amount of data written by the 
application increases to 750~GB, there is a greater disparity between 
Lustre and in-memory (2.67~x slower, on average). Local disk 
performance, however, 
is still comparable to memory (1.38~x slower, on average). Despite local disk and 
Lustre both being in transient state, local disk encounters less contention 
than what would be found on Lustre.

@@

At 100 iterations, or 7,500~GB, Lustre can be found to be, on average, 3.82x 
slower than Spark in-memory. The page cache, on Zenith, occupies 20\% of 
available memory. Assuming all memory (187~GB) was available, only 37.4~GB of 
data could be held in a node's page cache at any given time. The 
slowdown experienced can be explained by the much less data could 
reside in the page cache at a given time, in contrast to 10 iterations, 
and therefore the effects of disk bandwidth are more significant in 
this application. At 100 iterations, the application was writing 500GB 
per node (7,500 GB / 15 nodes) and hence could not run on tmpfs.

@@

Increasing task duration ensured that all file systems had a comparable performance
(Fig.~\ref{fig:cputime}). Lustre, for instance, is approximately 1.01x slower
than Spark in-memory at a task duration of 320 seconds, whereas it is 
approximately 3.25x slower that Spark in-memory with 2.4 second tasks. This 
pattern corroborated our page-cache model which postulates that 
data movement costs will have little impact on compute-intensive tasks. The 
reasoning for this is that longer tasks give the page cache more time to flush 
between write to disk. The longer the delay, the less likely the page cache is 
to fill up when the concurrent data size written fits within the page cache.

@@

% Comment on the general trend
As can be seen in Fig.~\ref{fig:numchunks}, makespan decreases when the number 
of chunks increases. This is due to the fact that parallelism increases. With 
30 chunks, only 2 CPUs per node are actively working. At 125 chunks, this 
changes to a maximum of 9 CPUs per node, and at 750 chunks all 40 CPUs can be 
active. It can be noted that no significant speedup can be seen between 125 
chunks and 750 chunks. This is because out of the 50 chunks processed by each 
node, only 40 can be processed concurrently, thus requiring two sequential 
incrementation batches, and resulting in similar makespans as 125 
chunks.

@@

% Spark local vs nipype local: presumably increases cache hits

@@

@@

@@

% fault-tolerance: if you loose a node, you loose the data when it's written on disk or in memory.
%                  then the only solution is to recompute, which spark would do automatically.
%                  when nipype is run on lustre, it would avoid recomputing existing files, which Spark doesn't do.
%                   this niype feature is lost when not using lustre.

>>>>

\begin{abstract}
    Neuroimaging datasets are rapidly growing in size as a result of 
    advancements in image acquisition methods, open-science and data sharing. 
    However, it remains that current neuroimaging data processing engines do 
    not employ strategies adopted by Big Data processing engines to improve 
    Big Data processing time. Here, we 
    evaluate three Big Data processing stategies (in-memory computing, 
    data-locality and lazy-evaluation) on typical neuroimaging use 
    cases, represented by the BigBrain dataset. We contrast the various 
    processing strategies by using Apache Spark and Nipype, as our 
    representative Big Data and neuroimaging processing engines, on Dell EMC's 
    Top-500 cluster. 
    Big Data thresholds were modeled by comparing the data/compute ratio of the 
    application to the filesystem/concurrent processes ratio. 
    This model acknowledges the 
fact that page caching provided by the Linux kernel is critical to the 
performance of Big Data applications. Results show that in-memory 
computing alone speeds-up executions by a factor of up to 1.6, whereas 
when combined with data locality, this factor reaches 5.3. Such important 
speed-up values are likely to be observed on typical image processing 
operations performed on images of size larger than 75GB. A ballpark 
speculation from our model showed that in-memory computing alone will
not speed-up current 
functional MRI analyses unless coupled with data 
locality, with a number of concurrently processed subject in the  200 ballpark. 
    In addition, we observe that emulating in-memory computing 
using in-memory file systems (tmpfs) does not reach the performance of 
an in-memory engine, presumably due to swapping to disk and the lack of data cleanup.
We conclude that Big Data processing strategies are 
worth developing for neuroimaging 
applications. 
\end{abstract}

@@

Big Data processing engines reduce data-related overheads through 
use of concepts such as data locality, in-memory computing, and lazy evaluation.
Data locality-aware scheduling, popularized by the MapReduce~\cite{dean2008mapreduce} 
framework, executes tasks in the nodes holding the data rather than 
performing costly data transfers across the network. In-memory 
computing, adopted by MapReduce's successor Apache 
Spark~\cite{zaharia2016apache}, stores data in memory whenever 
possible, reducing costly disk I/Os between pipeline steps. 
Lazy evaluation, also used in Spark, enables further Big Data performance 
optimizations, by evaluating results only when necessary. Frameworks 
such as
MapReduce and Spark have become mainstream tools for data analytics, 
although many others, such as 
Dask~\cite{rocklin2015dask}, are emerging. 
Meanwhile, several scientific 
domains including bioinformatics, physics or astronomy, have entered 
the Big Data era due to increasing data volumes and variety. 
Nevertheless, the adoption of Big Data engines for scientific data analysis 
remains limited, perhaps due to the widespread availability of 
scientific processing engines such as Pegasus~\cite{deelman2005pegasus} or
Taverna~\cite{oinn2004taverna}, and the adaptations required in Big 
Data processing engines for scientific computing.

Scientific applications differ from typical Big Data use 
cases, which might explain the remaining gap between Big Data and 
scientific engines. While Big Data applications mostly target text 
processing (e.g. web search, frequent pattern mining, recommender 
systems~\cite{leskovec2014mining}) implemented in consistent software 
libraries, scientific applications often involve 
binary data such as images and signals, processed by a sequence of 
command-line/containerized tools
using a mix of programming languages (C, Fortran, Python, shell 
scripts), referred to as workflows or pipelines. Infrastructure-wise, 
Big Data applications commonly run on 
clouds or dedicated commodity clusters with locality-aware file systems 
such as the Hadoop Distributed File System 
(HDFS~\cite{shvachko2010hadoop}), whereas scientific applications are 
usually deployed on large, shared clusters where data is moved between
data nodes and compute nodes using shared file systems such 
as Lustre~\cite{schwan2003lustre}. Such differences in applications and 
infrastructures have important consequences. To 
mention only one, in-memory computing requires instrumentation to be 
applied to command-line tools.

Technological advances of the past decade, in particular page caching 
in the Linux kernel~\cite{love2010linux}, in-memory file systems 
(tmpfs) and memory-mapped files might also 
explain the lack of adoption of Big Data engines for scientific 
applications. In such configurations, in-memory computing would be a feature 
provided by 
the operating system rather than by the engine itself. The frontier 
between these two components is blurred and needs to be clarified.

@@

Our study focuses on performance. We intentionally do not 
compare Big Data and scientific data processing engines on the grounds 
of workflow language expressivity, fault-tolerance, provenance capture 
and representation, portability or reproducibility which are otherwise 
critical concerns, addressed for instance in~\cite{samba}. Besides, our 
study of performance focuses on the impact of data writes and 
transfers. It purposely leaves out task scheduling to computing 
resources, to focus on the understanding of 
data writes and movement. Task scheduling will be part of our 
discussion, however.

@@

%\todo{Perhaps talk about modeling and simulation here.}

@@

An RDD is an immutable parallel data structure that achieves fault-tolerance 
through the concept of lineage~\cite{zaharia2010spark}. Rather than permitting
fine-grained transformations, only coarse-grained transformations (applying to
many elements in the RDD), can be applied, thereby making it simple to maintain a 
log of data modifications. This log, known as the lineage, is used
to reproduce any lost data.

Two types of operations can be performed on RDDs:
transformations and actions. Applying a transformation to an RDD produces a new
child RDD through a narrow or wide dependency. A narrow dependency signifies 
that the child is only dependent on a single parent partition, whereas a child 
RDD is dependent on all parent partitions in a wide dependency. Examples of 
Spark transformations include map, filter and join. To materialize an RDD, an
action must be performed, such as a reduce or a collect. Lazy evaluation is 
represented in Spark through the use of transformations and actions. A series of
transformations may be defined without the data ever being materialized. Using 
this strategy, Spark can make optimizations to ensure optimal processing speeds 
throughout the application.

%\todo{Lazy evaluation should be mentioned here, or elsewhere in the paragraph.
%It is currently not described.}

All actions and wide dependencies require a shuffle -- Spark's most costly
operation. Every shuffle begins with each map task saving its data to local
files for fault tolerance. The shuffle operation then 
redistributes the data across the partitions as requested. A shuffle marks a 
stage boundary in Spark. In other words, an reduce operation  will not begin 
until all 
dependent map tasks have completed.

Spark will spill RDD data to disk whenever the complete dataset does not fit in
memory. Moreover, as Spark transformations generate new RDDs and numerous 
transformations may occur in a single application, Spark implements a Least-Recently Used (LRU) eviction
policy. If an evicted RDD needs to be reused, Spark will recompute it using the
lineage data collected. As demonstrated in~\cite{freeman2014mapping
}, caching significantly improves processing times of iterative algorithms where 
RDDs are reused. It can be of even greater importance if the RDD is costly to 
recompute.
%mention results in thunder paper

Data locality in Spark is achieved through the scheduling of tasks to partitions 
which have the
data loaded in memory. If the data is instead stored on HDFS, the scheduler will
assign it to one of the preferred locations specified by HDFS. Spark's scheduler
utilizes delay scheduling to optimize fairness and locality for all tasks.
%improve paragraph

@@

Executing Spark applications on HPC systems with Spark-unaware schedulers may
be inefficient. The amount of resources requested by Spark may impede 
Spark-cluster scheduling time. Using pilot-scheduling strategies to add 
nodes to the Spark cluster as they are allocated by the underlying 
schedulers may speedup allocation and overall processing time~\cite{paraskevakos2018pilot}. This, however, is not studied in the 
current paper.

@@

% i use 'popular' for both...
Nipype is a popular Python neuroimaging processing engine. It aims at being a 
solution for easily creating reproducible neuroimaging workflows. Although Nipype
does not employ any Big Data processing strategies, it provides access to numerous
plugins to schedulers found in most clusters readily available to researchers, 
such as the SunGrid Engine (SGE), Portable Batch System (PBS), Slurm and HTCondor\todo{cite}. It also includes its own scheduler, MultiProc,
for parallel processing on single nodes. Nipype also includes many built-in 
interfaces to commonly used neuroimaging tools to be incorporated within the 
workflows. Nipype's ability to easily parallelize workflows in researcher-available 
cluster setups, capture detailed provenance information necessary for reproducibility,
and allow users to easily integrate existing neuroimaging tools, make it preferable
to existing Big Data solutions, which would necessitate modification to achieve this.

@@

In order for Nipype to operate as intended, a filesystem shared by all the nodes 
is required. However, it is still possible to save to a non-shared local filesystem, 
but it may come at the expense of fault-tolerance, as a results of being stored in a disk-local fs. Moreover, the user will need to
ensure that the files are appropriately directed to the nodes that require them as
there is no guarantee of data locality in Nipype.

@@

%\subsection{Simulation} % is it included? 0.5 page

% \todo{Remove if not included}

@@

 All experiments were executed on  
 \href{https://www.dellemc.com/resources/en-us/asset/sales-documents/products/storage/h16221-hpc-lab-brochure.pdf}{Dell EMC's Zenith 
 cluster}, a Top-500 machine in the Dell EMC HPC Innovation Lab, running 
 Slurm. For the Spark experiments, a Spark cluster was started on our 
 Slurm allocation through access to 16 dedicated nodes. Each compute 
 node ran on RHEL 7.4, was 
 equipped with 40 Intel Xeon Gold 6148 2.4GHz CPUs and had
 187GB of memory, 94GB of tmpfs and a Dell 120GB M.2 SATA SSD as local disk. 
 A 698TB Lustre server was accessible on each node through a 100Gb/s 
 Infiniband network. The apparent write bandwidth of the local disk, 
 Lustre file system and tmpfs were measured by sequentially writing various numbers
 of image blocks containing random intensities, to avoid 
 caching effects (measure\_bandwidth.py script in folder 
 benchmark\_scripts). They are reported in Table~\ref{table:bdwdths}.

@@

% Incrementation (avoid cache effects in binarization)
In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consisted exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the applications. As incrementing the images is a relatively quick 
process, we added a sleep delay to the incrementation stage in order to 
increase task duration as required. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the incremented chunks would be incremented again and saved to the same storage. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks and final output chunks would be read from / written to  Lustre and the final. 
We chose to perform our initial 
reads and final writes on Lustre as a multi-tenant clusters' compute node's 
local storage, if 
available, is generally temporary and only available for the duration of the 
allocation.

@@

We conducted four experiments in which we varied (1) the number of 
iterations in the application ($n$ in 
Algorithm~\ref{alg:incrementation}), (2) the task duration ($x$), (3) 
the chunk size for a constant image size, (4) the total image size. To 
evaluate the page-cache model, experiment conditions fell in different 
regions of Equations~\ref{eq:page-cache-inequality} 
and~\ref{eq:page-cache-sharedfs}, as summarized in 
Table~\ref{table:experiments}. Among the 16 nodes available, 1 was 
dedicated to the Spark master and driver, and the remaining 15 were 
used as compute nodes. Since data-locality is not normally preserved in 
Nipype (a new Slurm allocation is requested for each processed task), 
we instrumented Nipype to ensure data locality. That is, the chunk were split into partitions, and for each partition of 
chunks, we requested a Slurm allocation to process the entire 
pipeline in parallel, using Nipype's MultiProc scheduler, on a given 
node. This was possible as no communication was required between the 
processed chunks.

@@

As a third incrementation experiment, we were interested in the effects of 
chunk size on Big Data strategies. Naturally, greater chunk size signifies 
a decrease in parallelization, however, it also signifies an increase in 
sequential I/O (increased $\Delta/\Gamma$ and $\delta/\gamma$). For 
this experiment we partitioned the complete BigBrain image into 30, 125 
and 750 chunks, corresponding to $\gamma$ values of 2, 9 and 25 
respectively. While Spark attempted to load-balance the data, it used 
up only 25 of the 40 cores for 750 chunks. In contrast, Nipype tried to 
use up as many cores as possible. Unlike the previous 
experiment, the D/C ratio was kept static at 178.6MB/s, however, this 
ratio ensured that different regions of the inequality were reached 
depending on amount of parallelization. Unfortunately, 30 chunks could 
no be processed using in-memory Spark as its individual chunk size of 
3GB exceeded the 2GB partition size limit imposed by Spark. The number of iterations was
fixed to 10, and the task duration was adjusted so that C remained constant at 4,400s.

@@

Fig.~\ref{fig:iterations} shows the difference between the different 
filesystem choices given the number of iterations. At 1 iteration, all 
filesystems behave the same, although the application was writing faster than the page cache flushed.
This is explained by the fact that the application data was not large enough to saturate the page  
cache fully (transient phase). As the total amount of data written by the 
application increases to 750~GB, there is a greater disparity between 
Lustre and in-memory (2.67~x slower, on average). Local disk 
performance, however, 
is still comparable to memory (1.38~x slower, on average). Despite local disk and 
Lustre both being in transient state, local disk encounters less contention 
than what would be found on Lustre.

@@

At 100 iterations, or 7,500~GB, Lustre can be found to be, on average, 3.82~x 
slower than Spark in-memory. The page cache, on Zenith, occupies 20\% of 
total memory. Since there was 187~GB of RAM on each node, only 37.4~GB of 
dirty data could be held in a node's page cache at any given time. The 
slowdown experienced can be explained by the smaller percentage of total data 
residing in the page cache at a given time, compared to 10 iterations. 
Therefore, the effects of Lustre bandwidth are more significant in 
this application. At 100 iterations, the application was writing 500GB 
per node (7,500 GB / 15 nodes) and hence could not run on tmpfs or local disk.

@@

Increasing task duration ensured that all file systems had a comparable performance
(Fig.~\ref{fig:cputime}). Lustre, for instance, is approximately 1.01x slower
than Spark in-memory at a task duration of 320 seconds, whereas it is 
approximately 3.25x slower that Spark in-memory with 2.4 second tasks. This 
pattern corroborated our page-cache model which postulates that 
data movement costs will have little impact on compute-intensive tasks. The 
reasoning behind this is that longer tasks give the page cache more time to flush 
between write to disk. The longer the delay, the less likely the page cache is 
to fill up when the concurrent data size written fits within the page cache.

@@

% Comment on the general trend
As can be seen in Fig.~\ref{fig:numchunks}, makespan decreases when the number 
of chunks increases. This is due to the fact that parallelism increases with a increase in chunk. With 
30 chunks, only 2 CPUs per node are actively working. At 125 chunks, this 
changes to a maximum of 9 CPUs per node, and at 750 chunks all 40 CPUs can be 
active. It can be noted that no significant speedup can be seen between 125 
chunks and 750 chunks. This is because out of the 50 chunks processed by each 
node, only 40 can be processed concurrently, thus requiring two sequential 
incrementation batches and resulting in similar makespans as 125 
chunks.

@@

The effects of lazy-evaluation can be seen throughout the experiments. Nipype 
was found to generally be slower than Spark in all experiments. While a Nipype DAG is 
generated prior to workflow execution, there are no such optimizations, as in 
Spark, to ensure that the least amount of work is performed to produce the 
results.

During the processing of Experiment 3, it was found that Spark split up the data
into an processed it in two batches for the 750 chunk input. That is to say, 
each node in the cluster processed 50 chunks, and since there was only a maximum
of 40 cores available on each node, Spark opted to run to batches of 25 to ensure
a balanced workload. Rather than running each iteration on the full dataset, as with Nipype, 
Spark opted to perform all the iterations on the first batch (load, increment, save),
and then proceeds to process the second batch. Such an optimization is important, 
even when processing data on disk, as it would presumably increase the occurrance
of cache hits. This may partially explain the speedup seen at 750 chunks in
Figure~\ref{fig:numchunks}.% Spark local vs nipype local: presumably increases cache hits

@@

Although tmpfs and page cache do improve performance in general, as seen in
Figure~\ref{fig:results}, they do not always perform equivalently to in-memory. 
Tmpfs's main limitation is that data residing on it may be swapped to disk if 
the system's memory usage is high. When it reaches this point, its performance 
slows down to swap disk bandwidth, as observed in Figure~\ref{fig:datasize}. Page cache suffers a similar dilemma. As detailed
in~\href{https://access.redhat.com/documentation/en-us/red_hat_enterprise_linux/6/html/performance_tuning_guide/s-memory-tunables}{RedHatDoc}\todo{add to references},
I/O blocking write to disk occurs when a giving percentage (e.g. 20~\%) of total memory is occupied by dirty data, once this 
threshold is exceeded, processes wanting to execute writes must wait until the dirty data
in the cache is flushed to disk. Furthermore, like memory, tmpfs and page cache are shared resources on a node.
If users on the node are heavily using memory to incite tmpfs to write to swap space, or are performing data-intensive operations
that fill up swap space, this will limit the performance on tmpfs/page cache of other users. 
However, with memory, it is possible to request, through the HPC scheduler, a required
amount of available memory. Ultimately, in-memory data will also need to be spilled to
disk if memory usage exceeds amount of available memory, although disk writes are 
likely to occur in tmpfs and page cache before requested available memory is filled.

@@

An alternative Nipype plugin designed for running on Slurm was not used in the experiments.
The Slurm plugin requests a Slurm allocation for each processed data chunk. Such a
scheduling strategy was not ideal in our environment where oversubscription of nodes was 
not enabled.

Unlike Spark, Nipype, by default, opts to use all available CPUs rather than to load balance
data across the cluster. That is, given 50 chunks and 40 cores, Spark will only use up 25
cores, and will break it up into two batches. Nipype, on the other hand, will also have no 
choice but to split up the processing into two batches, but will first processed 40 chunks, 
immediately followed by the remaining 10. While both are reasonable strategies for 
data processing, Spark may end up benefiting more from the page cache, as less data is 
written in parallel (25 vs 40), giving more time for the page cache to flush.

@@

Writing to node-local storage in a cluster environment comes at a cost, for both
Nipype and Spark without HDFS.
% fault-tolerance: if you loose a node, you loose the data when it's written on disk or in memory.
%                  then the only solution is to recompute, which spark would do automatically.
%                  when nipype is run on lustre, it would avoid recomputing existing files, which Spark doesn't do.
%                   this niype feature is lost when not using lustre.
